---
title: "Class Assignment 1"
author: "Nahid Ferdous"
output: pdf_document
---

# Problem 1

Use the TTU graduate student exit survey data.

```{r tidy=FALSE}
grad <- read.csv("https://raw.githubusercontent.com/asheikhz2/TTU_Zadeh/main/pgs.csv")
```

a)  Create a new data-frame for three variables: "GenRating", "DeptStaff", "Housing".

```{r}
mydata = grad[, c("GenRating", "DeptStaff", "Housing")]

head(mydata)

```

b)  There are some missing values in this data. Find a correlation matrix for the data of part (a). If there are NAs (missing values) in your data, estimate the correlation matrix by all three following methods: (1) Complete Case Analysis, (2) MLE, (3) Median insertion.

```{r}
# Complete Case Analysis 
mydata_na_omit <- na.omit(mydata)
corr_na_omit <- cor(mydata_na_omit)
print("Complete Case Analysis")
corr_na_omit

# Maximum Likelihood Estimation
library(mvnmle)
mydata_MLE_fit <- mlest(mydata)
mydata_MLE_cov <- mydata_MLE_fit$sigmahat
mydata_MLE_corr <- cov2cor(mydata_MLE_cov)
cat("\n\n")
print("Maximum Likelihood Estimation")
mydata_MLE_corr

# Median insertion

mydata_Median_insertion <- mydata

for (c in 1: ncol(mydata_Median_insertion)){
  NaN_bool <- is.na(mydata_Median_insertion[,c])
  NaN_index <- which(NaN_bool)
  mydata_Median_insertion[NaN_index,c] <- median(mydata_Median_insertion[,c], na.rm = TRUE)
}
cat("\n\n")
print("Median insertion")
mydata_Median_insertion_corr = cor(mydata_Median_insertion)
mydata_Median_insertion_corr


```

# Problem 2

Read the crime data set.

```{r tidy=FALSE}
crime <- read.csv("https://raw.githubusercontent.com/asheikhz2/TTU_Zadeh/main/crime.csv", 
row.names = "STATE")
head(crime)
```

a)  Find the correlation matrix of the data.

```{r}
crime_cor <- cor(crime)
crime_cor
```

b)  Create a bi-variate boxplot for "MURDER" vs "AUTO".

```{r}
library(MVA)
MURDER_AUTO_data= crime[,c("MURDER", "AUTO")]
bvbox(MURDER_AUTO_data, xlab = "MURDER", ylab = "AUTO", col = "blue", cex = 0.5)
text(MURDER_AUTO_data, labels = row.names(crime), cex = 0.7, pos = 2)
```

c)  Create Chi-sq plot for mahalanobis distances (Lecture 8).

```{r}
x <- MURDER_AUTO_data
xbar <- colMeans(x)
s <- cov(x)
# mahalanobis distances 
d2 <- mahalanobis(x,xbar, s)
number_of_variables <- ncol(x)
position <- (1:nrow(x) -.5)/nrow(x)
quantiles <- qchisq(position, df = number_of_variables)

# chi_sq plot
plot(quantiles, sort(d2),
     xlab = expression(paste(chi[2]^2, "Quantile" )),
     ylab = "Ordered squared distances")
abline(a= 0, b =1)

text(quantiles, sort(d2), labels = names(sort(d2)), col = "red", pos = 2)
```

d)  Given what you see in part b and c, identify outliers. Remove those outliers, then find the correlation matrix. Compare the correlation matrices before and after removing outliers.

```{r}
# In Part B, we identified two outliers: 'RHODE ISLAND' and 'MASSACHUSETTS'. However, in Part C, only 'MASSACHUSETTS' was clearly an outlier, while 'RHODE ISLAND' wasn't definitively so.

print("Original correlation matrix")
crime_cor

cat("\n\n")
# first consider "RHODE ISLAND" and "MASSACHUSETTS" our outliers
two_outliers <- c("RHODE ISLAND", "MASSACHUSETTS")
two_outliers_indexs <- match(two_outliers, row.names(crime))
two_outliers_clean <- crime[-two_outliers_indexs,]
two_outliers_clean_corr <- cor(two_outliers_clean)
print("Two Outliers removed - RHODE ISLAND and  MASSACHUSETTS (correlation matrix)")
two_outliers_clean_corr

cat("\n\n")
# consider "MASSACHUSETTS" our outlier
one_outliers <- c("MASSACHUSETTS")
one_outliers_indexs <- match(one_outliers, row.names(crime))
one_outliers_clean <- crime[-one_outliers_indexs,]
one_outliers_clean_corr <- cor(one_outliers_clean)
print("One Outlier removed - MASSACHUSETTS (correlation matrix)")
one_outliers_clean_corr

cat("\n\n")
# Lets calculate Mean squared error
print("Original vs two outliers clean data ")
MSE_two_outliers_clean <- mean((two_outliers_clean_corr-crime_cor)^2)
print("MSE- Original vs Two_outliers_clean data")
MSE_two_outliers_clean

cat("\n\n")
print("Original vs one outlier clean data ")
MSE_one_outliers_clean <- mean((one_outliers_clean_corr-crime_cor)^2)
print("MSE- Original vs One_outliers_clean data")
MSE_one_outliers_clean

```

# Removing only 'MASSACHUSETTS' as an outlier results in a significantly lower MSE, indicating a better fit for the data compared to the model that also excludes 'RHODE ISLAND.' Therefore, 'RHODE ISLAND' may not be a true statistical outlier.
