---
title: "Homework 01- Problem 1"
author: "Nahid Ferdous"
date: "`r Sys.Date()`"
output: word_document
---
## a. For the 100-meter dash and 200-meter dash, compute the following summary statistics: mean, median, variance, standard deviation.
```{r }
track <- read.csv("https://raw.githubusercontent.com/EricBrownTTU/ISQS5346/main/menstrack.csv")
data_100m <- track[,1]
data_200m <- track[,2]

print("The summary statistics of 100m data: ")
print(paste("Mean : ", mean(data_100m )))
print(paste("Median : ", median(data_100m )))
print(paste("Variance : ", var(data_100m )))
print(paste("Standard deviation : ", sd(data_100m )))

print("The summary statistics of 200m data: ")
print(paste("Mean : ", mean(data_200m )))
print(paste("Median : ", median(data_200m )))
print(paste("Variance : ", var(data_200m )))
print(paste("Standard deviation : ", sd(data_200m )))

```

## b. Does does the use of the mode on the above data provide significant analytical value? Why or why not?
## Answer: No, mode is not perfect for continus data 

## c. Calculate the z-score for each of the observations in each of the two variables above. Which observations are the most extreme? Should we consider these values outliers?
```{r}
track <- read.csv("https://raw.githubusercontent.com/EricBrownTTU/ISQS5346/main/menstrack.csv")
z_scores <- scale(track[, c("m100", "m200")])
most_extreme <- track[which(abs(z_scores) == max(abs(z_scores)), arr.ind = TRUE), c("m100", "m200")]
most_extreme
```
## I think i dont need to consider these values as a outliers. 

## d. Create a dataset consisting of only the two variables above. Use cbind to incorporate a new variable corresponding to the 3000-meter run. Now with three variables, use the cor function to create a correlation matrix, which displays the correlations between each pairing of the three variables simultaneously. Comment on the relationships between the three events. Are they what you would expect?

```{r}

track <- read.csv("https://raw.githubusercontent.com/EricBrownTTU/ISQS5346/main/menstrack.csv")
sub_data <- track[,c("m100","m200")]
new_data= cbind(sub_data,track$m3000)

cor_matrix= cor(new_data)
cor_matrix

```

## The correlation matrix indicates a strong correlation between the 100-meter dash and 200-meter dash variables, suggesting that performance in the 100-meter event is generally associated with performance in the 200-meter event. This implies that athletes who excel in the 100-meter dash are likely to perform well in the 200-meter dash as well.

## However, the correlation between the 100-meter/200-meter events and the 3000-meter run is not strong. This suggests that the skills and abilities required for success in the 3000-meter run are different from those needed for the shorter sprints. Athletes who excel in the 100-meter and 200-meter events may not necessarily perform well in the 3000-meter run, indicating a distinction between the two groups of athletes.

## In summary, the correlation matrix suggests a strong relationship between the 100-meter and 200-meter events, indicating similar performance levels. However, the lack of a strong correlation with the 3000-meter run suggests that athletes specializing in longer distances have different characteristics and abilities compared to those in the shorter sprints. 
