---
title: "Homework_2_Problem_2"
author: "Nahid Ferdous"
date: "`r Sys.Date()`"
output: word_document
---
## a. What is the expected value and standard deviation of the number of patrons in your restaurant on a given day? Hint: Assume that ğ‘‹ is a Poisson-distributed random variable. What is ğœ†?
## The expected value (mean) of the Poisson distribution is given by ğ¸(ğ‘‹) = ğœ†, and the standard deviation is given by ğœ(ğ‘‹) = sqrt(ğœ†).

## Given that you see around 600 patrons on an average day, ğœ† would be equal to 600.

## Therefore, the expected value of the number of patrons is ğ¸(ğ‘‹) = 600, and the standard deviation is ğœ(ğ‘‹) = sqrt(600).
```{r}
ğœ†= 600
Expacted_value= ğœ†
Expacted_value
standard_deviation = sqrt(ğœ†)
standard_deviation

```
## b. Find the possibility of all possible values of ğ‘‹ graphically and the collective probability of each grouping of one hundred patrons (e.g. ğ‘ƒ(0 â‰¤ ğ‘‹ < 100), ğ‘ƒ(100 â‰¤ ğ‘‹ < 200) etc.) numerically. 
```{r}
x <- 0:1200
lambda <- 600
probabilities <- dpois(x, lambda)

plot(x, probabilities, type = "h",
     main = "Probability Distribution of X",
     xlab = "Number of Patrons (X)",
     ylab = "Probability",
     col = "blue"
)

```
## To calculate the collective probability of each grouping of one hundred patrons (e.g., ğ‘ƒ(0 â‰¤ ğ‘‹ < 100), ğ‘ƒ(100 â‰¤ ğ‘‹ < 200), etc.) numerically, we can use the ppois() function in R. The code snippet below calculates the probabilities for the specified groupings:
```{r}
groupings <- seq(0, 1200, 100)
group_probabilities <- diff(ppois(groupings, lambda))

# Display the probabilities for the groupings
for (i in 1:length(groupings)) {
  cat(paste0("P(", groupings[i], " â‰¤ X <", groupings[i] + 100, ") = ", group_probabilities[i], "\n"))
}


```
## c. Your cooks are getting cranky and have told you that they will quit if there are more than 675 patrons ever. What is the probability of such an event? 
```{r}
probability_more_than_675 <- 1 - ppois(675, lambda)
probability_more_than_675
```
## The probability_more_than_675 variable will contain the probability of having more than 675 patrons in your restaurant on a given day. 

## d. Simulate your restaurant workload for the next year (take a look at the function rpois()). Compare your results against those from part b. 
```{r}
num_days <- 365
simulated_workload <- rpois(num_days, lambda)

hist(simulated_workload, breaks = 20, 
     main = "Simulated Workload Distribution",
     xlab = "Number of Patrons",
     ylab = "Probability",
     col = "blue",
     border = "white",
     probability = TRUE
)

# Calculate the probabilities
normalized_freq <- hist(simulated_workload, plot = FALSE)$counts / num_days
normalized_freq

# Display the probabilities above each bar
mtext(text = round(normalized_freq, 4), side = 3, at = pretty(simulated_workload), line = -2, cex = 0.7)

```
