---
title: "Untitled"
author: "Nahid Ferdous"
date: "`r Sys.Date()`"
output: word_document
---

```{r}
panel <- read.csv("https://raw.githubusercontent.com/EricBrownTTU/ISQS5346/main/panel.csv")
panel[,1]
mean(panel[,1])
median(panel[,1])
sd(panel[,1])
```
```{r}
qqnorm(panel[,1])
qqline(panel[,1])
```
## Most of the data fall in accordance with the theoretical quantiles. We can now look at the skewness and kurtosis.

```{r}
x<- panel[,1]
z<- (x-mean(x)/sd(x))
k3= sum(z^3)/length(z)
k3

k4= sum(z^4)/length(z)-3
k4
```

## Both the skewness and kurtosis are within acceptable limits, so we must conclude that the data that we have is sufficiently close to normal.

# Rankings of Universities
## Consider the data on academic reputation scores for the top 50 research universities.

```{r}
topuniv <- read.csv("https://raw.githubusercontent.com/EricBrownTTU/ISQS5346/main/topuniv.csv")
x <- topuniv[,4]
qqnorm(x)
qqline(x)

```

## The data do not adhere well to the linear relationship between quantiles.


```{r}
z <- (x - mean(x))/sd(x)
K3 = sum(z^3)/length(z)
K3
## [1] 0.1411848
K4 = sum(z^4)/length(z) - 3
K4
## [1] -1.183137
```
## The skewness and kurtosis values are reasonable, which indicates the data might be normal. I would argue that this data requires additional inspection before modeling using the normal distribution.