---
title: "R Practice"
author: "Nahid Ferdous"
date: "`r Sys.Date()`"
output: word_document
---

```{r }
x <- read.csv("https://raw.githubusercontent.com/EricBrownTTU/ISQS5346/main/measure.csv") 
head(x)
```

## Input all data.

## Leats standarized the data

```{r}
scale(x[,1:3])
```

## Here scale() is a cool function, this function bassically convert the variables value
## to z score.

## Marginal and Conditional Distributions 
## Functions for marginal and conditional distributions: "MASS"

```{r}
library(MASS)
head(Cars93)

```

```{r}
xt<- table(Cars93$Type,Cars93$Origin)
xt
colSums(xt)
rowSums(xt)


```

## first lets constant collum variables so marzin=01 
## second lets consider the row in constant 
## here : CR = 1,2 (1= collums, 2= rows)

```{r}
prop.table(xt,margin=1)
prop.table(xt,margin=2)
```
## For Cramer's v we need to install a new package , the name of the package is "Recompanion" 
```{r}
library(rcompanion)
v <- cramerV(xt)
v
```

## Again Covariance and Correletions 

```{r}
print("Covariance : ")
cov(x[,1:3])
print("Correletions: ")
cor(x[,1:3])
```
